

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
library(magrittr)
library(dplyr)
library(readr)
library(tidyr)
library(lubridate)
library(pander)
library(ggplot2)
library(grid)
library(gridExtra)
library(cowplot)
library(stringr)
library(ggmap)
```


# Loading the Data

```{r load_data}
#confirmed COVID-19 time series cases US (trimmed to include from 01/01/2021 to 05/23/2022)
covid.usa.ts.confirmed <- read_csv('https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_US.csv') %>%  select(UID:Combined_Key,`1/1/21`:`5/23/22`)

#Confirmed COVID-19 time series deaths US (trimmed to include from 01/01/2021 to 05/23/2022) these data include the population by county
covid.usa.ts.deaths <- read_csv('https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_US.csv') %>%
  select(UID:Population,`1/1/21`:`5/23/22`)

#Daily data summary by state for 05-23-20222
covid.usa.daily <- read_csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports_us/05-23-2022.csv") 

#US vaccinated people data
vacc.people <- read_csv("https://raw.githubusercontent.com/govex/COVID-19/master/data_tables/vaccine_data/us_data/time_series/people_vaccinated_us_timeline.csv")
```


# Data Manipulation

Combining the three data sets into a single tibble with columns of State, Date, Confirmed Cases, Confirmed Deaths, People Partially Vaccinated, and People fully Vaccinated.

```{r}

#Reshaping with pivot_longer and trimming to state, date, and cases/deaths columns

#covid.usa.ts.confirmed

my.ts.confirmed <- covid.usa.ts.confirmed %>% 
  pivot_longer(cols = contains("/"),names_to = "Date", values_to = "confirmed_cases") %>% 
  select(Province_State, Date, confirmed_cases) %>% 
  group_by(Province_State, Date) %>%
  summarise(confirmed_cases = sum(confirmed_cases)) %>% 
  ungroup() %>% 
  mutate(Date = mdy(Date)) %>% 
  arrange(Province_State, Date)
  
my.ts.deaths <- covid.usa.ts.deaths %>%
  pivot_longer(cols = contains("/"),names_to = "Date", values_to = "confirmed_deaths") %>% 
  select(Province_State, Date, confirmed_deaths) %>% 
  group_by(Province_State, Date) %>%
  summarise(confirmed_deaths = sum(confirmed_deaths)) %>% 
  ungroup() %>% 
  mutate(Date = mdy(Date)) %>% 
  arrange(Province_State, Date)

#my.ts.confirmed
#my.ts.deaths

#Creating covid.usa.states.ts

covid.usa.states.ts <- my.ts.confirmed %>%
   full_join(my.ts.deaths, by = c("Province_State", "Date"))

#covid.usa.states.ts


#Appending covid.usa.states.ts with vacc.people

#trimming down vacc.people
my.vacc.people <- vacc.people %>% 
  select(Province_State, Date, People_Fully_Vaccinated, People_Partially_Vaccinated) %>% 
  filter(!is.na(People_Fully_Vaccinated), !is.na(People_Partially_Vaccinated))

#my.vacc.people

#joining
covid.usa.states.ts <- covid.usa.states.ts %>% full_join(my.vacc.people, by = c("Province_State", "Date"))

covid.usa.states.ts

```

## Visualizing Covid Data from 6 select states

I will be selecting 3 states with the highest confirmed cases (California, Texas, and Florida) and 3 arbitrarily selected states with the mild amounts confirmed cases(West Virginia, Oregon, and Connecticut). These selections are based on the total number of cases as seen from the following table.

``` {r}

covid.usa.daily <- arrange(covid.usa.daily, Confirmed)
covid.usa.daily
```

Each variable will be standardized by dividing each variable by its maximum, so that they will all fit into the same graph and still be discernible. A ratio of 1 means a variable is at it's max on that day.

```{r}

 
# #scaling values by variable/maximum so all three variables can be overlayed on the same line graph and keeping only the 6 relevant states

#covid.usa.states.ts

adjusted.covid.usa.states.ts <- covid.usa.states.ts %>% 
  filter(Province_State %in% c('California', 'Texas', 'Florida', 'West Virginia', 'Oregon', 'Connecticut')) %>% 
  group_by(Province_State) %>%
  summarise(Date, confirmed_cases = confirmed_cases/max(confirmed_cases, na.rm = TRUE),
          confirmed_deaths = confirmed_deaths/max(confirmed_deaths, na.rm = TRUE),
          People_Fully_Vaccinated = People_Fully_Vaccinated/max(People_Fully_Vaccinated, na.rm = TRUE),
          People_Partially_Vaccinated = People_Partially_Vaccinated/max(People_Partially_Vaccinated, na.rm = TRUE))

#adjusted.covid.usa.states.ts

ggplot(data = adjusted.covid.usa.states.ts, aes(x = Date))+
  geom_line(aes(y = confirmed_cases), col = "red")+
  geom_line(aes(y = confirmed_deaths), col = "blue")+
  geom_line(aes(y = People_Fully_Vaccinated), col = "green")+
  geom_line(aes(y = People_Partially_Vaccinated), col = "purple")+
  xlab("Date")+
  ylab("Standardized ratio")+
  #scale_color_manual(name = "Variables:", values = c("Confirmed Cases" = "red", "Confirmed Deaths" = "blue", "Fully Vaccinated" = "green", "Partially Vaccinated" = "purple")) +
  labs(title = "Standardized data by state")+
  facet_wrap(~Province_State)
```
``` {r, echo=FALSE}
#+
  # theme(legend.position = "right")+
  # scale_color_manual(name = "Variable", 
  #                    values = c("Confirmed Cases" = "red", 
  #                               "Confirmed Deaths" = "blue", 
  #                               "People Fully Vaccinated" = "green",
  #                               "People Partially Vacinated" = "purple"))
  


#making a separate legend because I can't get the ggplot one to work properly

names <- c('Confirmed Cases', 'Confirmed Deaths', 'People Fully Vaccinated', 'People Partially Vaccinated')
clrs <- c('red', 'blue', 'green', 'purple')
ltype <- c(1, 1, 1, 1)
plot(NULL ,xaxt='n',yaxt='n',bty='n',ylab='',xlab='', xlim=0:1, ylim=0:1)
legend("topleft", title="Variables", legend = names, lty=ltype, lwd=2, cex=1.25,
       bty='n', col = clrs)


  #scale_color_manual(name = "Variables:", values = c("Confirmed Cases" = "red", "Confirmed Deaths" = "blue", "Fully Vaccinated" = "green", "Partially Vaccinated" = "purple"))


```


## Creating a Map visualization of cases in Oregon:

How were the confirmed cases in Oregon Distributed on 4/26/21?
```{r}

loc.or.confirmed <- covid.usa.ts.confirmed %>% 
  pivot_longer(cols = contains("/"),names_to = "Date", values_to = "confirmed_cases") %>% 
  filter(Province_State == "Oregon", Date == "4/26/21", Lat != 0) %>% 
  select(Lat, Long_, confirmed_cases)

loc.or.confirmed

ggplot(loc.or.confirmed, aes(x = Long_, y = Lat, col = confirmed_cases, size = confirmed_cases))+
  borders("state")+
  geom_point()+
  scale_color_viridis_c()+
  coord_quickmap(xlim = c(-125, -116), ylim = c(42, 46.5))+
  xlab("Longitude") +
  ylab("Latitude") +
  labs(title = "Oregon Confirmed Cases Distribution of 4/26/21")


```

